I would like to express my sincere gratitude to Professor Fabrizio Ferrandi for extending to me the wonderful opportunity to take part in this enlightening academic journey.
A deep sense of appreciation goes to my co-advisors, Serena Curzel and Michele Fiorito, for their mentorship and indispensable assistance.
Their patient guidance, profound expertise, and willingness to address my questions have played a crucial role in refining the trajectory of my work.
The insights they shared with me have not only enhanced the quality of my research but have also enlarged my knowledge, enabling me to approach challenges with a more informed perspective.

I extend my warmest thanks to my family, whose constant encouragement, support, and belief in my capabilities have been the foundation of my efforts.
Their faith in my potential has provided me with the strength to persist through challenges and to continually strive for excellence.
Lastly, I would like to express my gratitude to Caterina, a constant source of inspiration and unwavering support.
Her presence by my side has been a source of motivation, enabling me to face challenges with renewed determination.
Her belief in me has been a driving force that has propelled me forward, even during moments of uncertainty.

In conclusion, the collective support, guidance, and belief of these remarkable people have left an indelible imprint on both my academic pursuits and personal development.
Their contributions have not only facilitated the realization of my academic goals but have also enriched me with invaluable life lessons that I will carry with me throughout my journey.